\headingfont University of Extremadura\\ Faculty of Science\\[1cm] \textbf{ Physics degree\\ Degree Final Project\\[3cm] \flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm] }

### 2020-06-11 docs/paper/LaTex/Paper.tex
Checking bibliography references and completing docs

This is more or less, the definition of the IoT (Internet of Things) concept; according to \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications. However, also according to \cite{IoTOverview} there is not an official definition for IoT concept, but according to \cite{IoTObjetives}

@@

Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way. This could turn out bothering because the big size for this statistical population, \lang{as it was exposed before}. So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.

@@

\subsubsection{LoRa (Long Range)}
LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite{LoRaGeneral}. This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed. So this low availability along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.

In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite{LoRaGeneral}

@@

By other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN). The former one (the physical layer), is a proprietary technology developed by Semtech. So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite{LoRaGeneral}. This particularly means that LoRaWAN protocol is perfect for the purpose of this project.

@@

\subsubsection{Dragino shields/hats}
These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages. Both shields are based on the SX1276/SX1278 transceiver. However the RaspberryPi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.

@@

\subsubsection{gateway}
Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role. These must be generic devices due LoRaWAN protocol versatility. LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol. So it is possible for any device which implements hardware for a LoRa physical layer, to act as a gateway.

Nevertheless, there are important considerations about the said above,

>>>>

This is more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications. However, also according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept, but according to \cite[p.~2(920)]{IoTObjetives}

@@

Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way. This could turn out bothering because the big size for this statistical population, \lang{¿as/like? it was exposed ¿before/above?}. So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.

@@

\subsubsection{LoRa (Long Range)}
LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}. This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed. So this low availability along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.

In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite[]{LoRaGeneral}

@@

By other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN). The former one (the physical layer), is a proprietary technology developed by Semtech. So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}. This particularly means that LoRaWAN protocol is perfect for the purpose of this project.

@@

\subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages. Both shields are based on the SX1276/SX1278 transceiver. However the RaspberryPi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.

@@

\subsubsection{Gateway}
Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role. These must be generic devices due LoRaWAN protocol versatility. LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol. So it is possible for any device which implements hardware for a LoRa physical layer, to act as a gateway.

Nevertheless, there are important considerations about the said above, for example, nodes are not actually associated with an specific gateway. Instead, data transmitted by a node is typically received by multiple gateways. Each gateway will forward the received packets from the end-node to de cloud-based network server. Besides, this project is using a RaspberryPi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because means, according to Semtech, this transceiver is not intended to play a gateway role, but a end-node role.

A transceiver, by definition, is a device that is able to both, transmit and receive (in fact, the word itself is a mix between both, \textbf{trans}miter and re\textbf{ceiver}) that is what a node must be able to do; i.e. transmit the sensor data (context) and receive data to perform operations with its actuators.

Nevertheless, LoRaWAN specification varies from region to region \enquote{based on
the different regional spectrum allocations and regulatory requirements}\cite[p.~12-13]{LoRaWANspec}
